Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network

### 2017-02-28 websci-2017/source/paper.tex
Update text.

\epigraph{The Net interprets censorship as damage and routes around it.}{\textit{--John Gilmore}\cite{first_elmer-dewitt_1993}}

Much of the world's infrastructure is networked: power grids,
cellular networks, raods, and of course, the Internet.
With any critical infrastructure, the cost of failures can be
immense, so methods for tolerating various kind of faults within networks are an
important and ongoing area of research
\cite{zin_survey_2015,albert_error_2000,sterbenz_resilience_2010}.
{\em Adversarial faults},
those in which an adversary can target attacks strategically,
deserve special attention.
Such attacks are both extremely difficult to guard against and 
often have important social implications.
In particular, censorship and surveillance are often achieved
by targeting central network locations and either blocking or capturing
the information flowing through them.
The Internet's decentralized design was motivated
by the need to withstand targeted attacks, such as nuclear strikes
\cite{baran_distributed_1964}.
But despite longstanding common wisdom \cite{first_elmer-dewitt_1993},
both theoretical results and recent events (described below)
have demonstrated that
the Internet is surprisingly vulnerable to attack.
Decentralization remains a promising approach towards
building resilient networks,
but further work is needed to understand the relationship between
decentralized network structure and adversarial fault tolerance.

The Internet's vulnerability to censorship and other targeted attacks
has been demonstrated by several recent events.
In 2008, YouTube suffered a worldwide outage for several hours
when a service provider in Pakistan advertised false routing information
\cite{hunter_pakistan_2008}.
The action (known as a {\em black hole attack}) was intended to censor YouTube
within Pakistan only, but resulted in a worldwide cascading failure.
The action was initiated by government order in Pakistan,
and spread beyond Pakistan when a single misconfigured router allowed the false
routing information to propagate globally.
While the government order and router misconfiguration initiated the outage,
it was the structure of the Internet's router network that allowed a fault in a
single router to propagate globally.
And while the action was not an intentional attack against the global Internet,
the ability of an attacker to succeed without trying only strengthens our
claim of the Intenrnet's vulnerability.

Such vulnerabilities are not limited to any one system or protocol,
but result from the network structure itself.

@@

With strong theoretical and historical evidence that network structure
can create vulnerabilities,
methods for analyzing structural vulnerabilities and for designing
fault-tolerant networks are needed.
This paper presents several contributions towards advancing those goals.

We consider a setting in which a source node (Alice)
in a network attempts to route a message
to a target node (Bob) by forwarding it through the links of the network.
A ``link'' in this context could represent any kind of connection
(e.g., physical cables, encrypted channels).
We assume that some nodes in the network may be compromised by an attacker
(Mal).
Compromised nodes may behave incorrectly by blocking, altering,
or incorrectly routing messages.
We assume that Mal has full knowledge of the network structure, but has
limited resources and thus can only compromise a fixed number of nodes.
We also assume that nodes within a fixed network
distance of either Alice or Bob are
{\em trusted}
and cannot be compromised.
This assumption,
which we call
{\em partial trust transitivity},
is a weaker (more realistic) form of the transitive
trust assumption used by the popular 
web of trust approach
\cite{zimmermann_official_1995,ferguson_practical_2003}.

Under the above assumptions,
we show how to evaluate the influence of network structure
on attack-tolerance,
and propose a method for achieving a high-level of attack-tolerance on
the butterfly network topology.
The butterfly topology is popular in parallel processing
\cite{kshemkalyani_distributed_2008} and
peer-to-peer \cite{lua_survey_2005, korzun_structured_2013}
applications, due to its regular structure and high connectivity.
While it may seem questionable to dictate the topology of trust relationships,
there are many applications where this occurs,
e.g., overlay networks
\cite{lua_survey_2005, korzun_structured_2013},
formal organizations \cite{mohr_explaining_1982},
government-regulated cellular networks \cite{walker_mass_2012},
and call tree notification systems \cite{nickerson_thinking_2010}.
Furthermore, we stress that trust is not static;
it changes over time (ableit slowly).
In this sense, we approach trust as deliberate infrastructure.
Since centralized network structure is inherently
vulnerable to attack, the ability to deliberately structure trust relationships
is a necessary component of any attack-tolerance scheme.

>>>>

\epigraph{The Net interprets censorship as damage and routes around it.}{\textit{--John Gilmore}\cite{elmer-dewitt_first_1993}}

Much of the world's infrastructure is networked: power grids,
cellular networks, raods, and of course, the Internet.
With any critical infrastructure, the cost of failures can be
immense, so methods for tolerating various kind of faults within networks are an
important and ongoing area of research
\cite{zin_survey_2015,albert_error_2000,sterbenz_resilience_2010}.
{\em Adversarial faults},
those in which an adversary can target attacks strategically,
deserve special attention.
Such attacks are both extremely difficult to guard against and 
often have important social implications.
In particular, censorship and surveillance are often achieved
by targeting central network locations and either blocking or capturing
the information flowing through them.
The Internet's decentralized design was motivated
by the need to withstand targeted attacks, such as nuclear strikes
\cite{baran_distributed_1964}.
But despite longstanding common wisdom \cite{elmer-dewitt_first_1993},
both theoretical results and recent events (described below)
have demonstrated that the Internet can be surprisingly vulnerable to attack.
Decentralization remains a promising approach towards
building resilient networks,
but there is a need to better understand the relationship between
decentralized network structure and adversarial fault tolerance.

The Internet's vulnerability to censorship and other targeted attacks
has been demonstrated by several recent events.
In 2008, YouTube suffered a worldwide outage for several hours
when a service provider in Pakistan advertised false routing information
\cite{hunter_pakistan_2008}.
The action (known as a {\em black hole attack}) was intended to censor YouTube
within Pakistan only, but resulted in a worldwide cascading failure.
The action was initiated by government order in Pakistan,
and spread beyond Pakistan when a single misconfigured router allowed the false
routing information to propagate globally.
While the government order and router misconfiguration initiated the outage,
it was the structure of the Internet's router network that allowed a fault in a
single router to propagate globally.
And while the action was not an intentional attack against the global Internet,
the ability of an attacker to succeed without trying only strengthens our
claim of the Intenrnet's vulnerability.
In 2013, the Texas-based email provider Lavabit was compelled by to turn
their private SSL keys over to the FBI \cite{poulsen_edward_2013}.
In order to protect their users from surveillance,
Lavabit ceased operations.
While originally intended as surveillance,
this action effectively became an act of censorship.
In this case, the attack was again successful due to a highly centralized
architecture:
SSL keys under control of a single entity, in a single legal jurisdiction.
It is also important to note that there was no technological failure,
the attack was achieved through coercion.
So we see that such vulnerabilities are not limited to any one system or protocol,
but result from centralized structure itself.

@@

With strong theoretical and historical evidence that centralized
network structure can create vulnerabilities,
methods for analyzing structural vulnerabilities and for designing
fault-tolerant networks are needed.
This paper presents several contributions towards advancing those goals.

We consider a setting in which a source node (Alice)
in a network attempts to route a message
to a target node (Bob) by forwarding it through the links of the network.
A ``link'' in this context could represent any kind of connection
(e.g., physical cables, encrypted channels).
We assume that some nodes in the network may be compromised by an attacker
(Mal).
We also assume that Mal is an adversary of Alice specifically and targets
nodes strategically with the goal of interfering with Alice's communications
(rather than disrupting the network as a whole).
Compromised nodes may behave incorrectly by blocking, altering,
or incorrectly routing messages.
We assume that Mal has full knowledge of the network structure, but has
limited resources and thus can only compromise a fixed number of nodes.

We also assume that nodes {\em trust} their immediate neighbors,
so Mal is unable to compromise Alice's node, or her direct neighbors.
In the commonly used {\em web of trust} approach
\cite{zimmermann_official_1995,ferguson_practical_2003},
we would extend that trust transitively to the entire network.
However, we make a weaker and more realistic assumption:
that trust is extended transitively to nodes within a fixed
distance.
So only the beginning and end of a path between Alice and Bob is trusted
against interference from Mal.
We call this assumption
{\em partial trust transitivity},
and refer to such paths as {\em partially trusted}.

Under the above assumptions,
we show how to evaluate the influence of network structure
on attack-tolerance,
how to constructed trusted channels from redundant partially trusted paths,
and propose a method for achieving a high-level of attack-tolerance on
the butterfly network topology.
The butterfly topology is popular in parallel processing
\cite{kshemkalyani_distributed_2008} and
peer-to-peer \cite{lua_survey_2005, korzun_structured_2013}
applications, due to its regular structure, low degree, and high connectivity.

The reader may wonder whether it is realistic or useful to assume control over
the network structure.
We have already seen that whenever a single point of failure exists in the network,
there is a potential to exploit it through coercion,
without an attacker needing to compromise the technology.
So, {\em attack-tolerance cannot be achieved without the ability to influence
network structure}.
Luckily, there are scenarios in which network topology can be dictated.
Examples include overlay networks
\cite{lua_survey_2005, korzun_structured_2013},
formal organizations \cite{mohr_explaining_1982},
government-regulated cellular networks \cite{walker_mass_2012},
and call tree notification systems \cite{nickerson_thinking_2010}.
In general,
{\em when the need for attack-tolerance is high enough to warrant investment
in infrastructure, networks can be engineered and maintained as infrastructure}.
It is also worth noting that attack-tolerant networks may be sub-components of
larger, less-constrained systems.
For example, a single server might be replaced by a distributed network of servers,
each with different ownership, physical location, and legal jurisdiction,
without placing any unrealistic constraints on the clients connecting to
those servers.
